---
title: "Capstone Project"
author: "John Karuitha"
date: "27/10/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Project Overview: MovieLens**
For this project, you will be creating a movie recommendation system using the MovieLens dataset. The version of movielens included in the dslabs package (which was used for some of the exercises in PH125.8x: Data Science: Machine Learning) is just a small subset of a much larger dataset with millions of ratings. You can find the entire latest MovieLens dataset here. You will be creating your own recommendation system using all the tools we have shown you throughout the courses in this series. We will use the 10M version of the MovieLens dataset to make the computation a little easier.

```{r}
setwd("C:\\Users\\johnk\\OneDrive - University of Witwatersrand\\r_training\\capstone")
load("C:/Users/johnk/OneDrive - University of Witwatersrand/r_training/capstone/data.RDA.RData")
save.image("C:/Users/johnk/OneDrive - University of Witwatersrand/r_training/capstone/data.RDA.RData")
```


You will download the MovieLens data and run code we will provide to generate your datasets.

First, there will be a short quiz on the MovieLens data. You can view this quiz as an opportunity to familiarize yourself with the data in order to prepare for your project submission.

Second, you will train a machine learning algorithm using the inputs in one subset to predict movie ratings in the validation set. Your project itself will be assessed by peer grading.

# **Introduction**
You will use the following code to generate your datasets. Develop your algorithm using the edx set. For a final test of your final algorithm, predict movie ratings in the validation set (the final hold-out test set) as if they were unknown. RMSE will be used to evaluate how close your predictions are to the true values in the validation set (the final hold-out test set).

Important: The validation data (the final hold-out test set) should NOT be used for training, developing, or selecting your algorithm and it should ONLY be used for evaluating the RMSE of your final algorithm. You should split the edx data into separate training and test sets to design and test your algorithm.

Also remember that by accessing this site, you are agreeing to the terms of the `edX Honor Code`. This means you are expected to submit your own work and can be removed from the course for substituting another student's work as your own.

# **Create train and validation sets**
```{r, eval FALSE}
##########################################################
# Create edx set, validation set (final hold-out test set)
##########################################################

# Note: this process could take a couple of minutes

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(data.table)

# MovieLens 10M dataset:
# https://grouplens.org/datasets/movielens/10m/
# http://files.grouplens.org/datasets/movielens/ml-10m.zip

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

# if using R 3.6 or earlier:
#movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                            #title = as.character(title),
                                            #genres = as.character(genres))
# if using R 4.0 or later:
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
                                            title = as.character(title),
                                            genres = as.character(genres))


movielens <- left_join(ratings, movies, by = "movieId")

# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding") # if using R 3.5 or earlier, use `set.seed(1)`
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
      semi_join(edx, by = "movieId") %>%
      semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

# **Explore the Data**
```{r}
head(edx)
head(validation)
```
# **Q1**
How many rows and columns are there in the edx dataset?

Number of rows:
Number of columns:

```{r}
nrow(edx)
ncol(edx)
```
# **Q2**
How many zeros were given as ratings in the edx dataset?
How many threes were given as ratings in the edx dataset?

```{r}
summary(edx$rating)
table(edx$rating)
```

# **Q3**
How many different movies are in the edx dataset?

```{r}
length(unique(edx$movieId)) ## Taken as correct answer
length(unique(edx$title))
```

# **Q4**
How many different users are in the edx dataset?

```{r}
length(unique(edx$userId))
```

# **Q5**
How many movie ratings are in each of the following genres in the edx dataset?

```{r}
edx %>% filter(str_detect(genres, "Drama")) %>% count(genres) %>% summarize(u = sum(n))
edx %>% filter(str_detect(genres, "Comedy")) %>% count(genres) %>% summarize(v = sum(n))
edx %>% filter(str_detect(genres, "Thriller")) %>% count(genres) %>% summarize(w = sum(n))
edx %>% filter(str_detect(genres, "Romance")) %>% count(genres) %>% summarize(x = sum(n))
```

Which movie has the greatest number of ratings?

```{r}
edx %>% group_by(title) %>% 
  
  count(rating) %>% summarize(k = sum(n)) %>% 
  
  filter(str_detect(title, "Forrest Gump|Jurassic Park|Pulp Fiction|The Shawshank Redemption|Speed 2: Cruise Control")) %>% 
  
  arrange(desc(k))
```

# **Q7**
What are the five most given ratings in order from most to least?

```{r}
sort(table(edx$rating), decreasing = TRUE)
```

